Apache Storm

### 2018-01-25 paper.tex
Minor changes to report as suggested

Data can be considered the oil of this present age, because with huge amounts of data one can determine hidden patterns. A lot of companies utilize pattern recognition on big data to open opportunities for revenue generation whereas research institutions use it to discover patterns simply not observable by the human mind(simply due to the volume of data). The question on how to classify any data as big data can be answered by using the 3Vs: Volume, Variety and Velocity \cite{usingapachestorm}. Any data with big volume or variety (such as text, video, images, etc) or that gets generated rapidly can be considered as big data. In this paper regarding \textbf{Apache Storm} we consider the third V of big data to be the influencing factor.

\textbf{Data Stream} is a sequence of data flowing from a source to a destination, for example a sensor reporting the temperature to a server. The sensor measures the temperature periodically and updates it to the server. Another example could also be the purchase requests of products in a shopping site. In all of these scenarios, the data is continuously being generated. Traditionally analyzing of these types of data requires storage of these data onto persistence and later batch processing it to discern patterns as well as other useful information. But, with the help of \textbf{Apache Storm} we are able to process these data as they are generated with minimum amount of latency.

@@

\paragraph{The Master node} also called the \textbf{Nimbus} is also responsible for distribution and coordination of the execution of topologies. The Nimbus is the first point where the user submits the defined topology, which is then distributed among the worker nodes.  Figure \ref{fig:arch} illustrates the high level architecture where, Nimbus monitors all the available worker nodes and both the nimbus as well as the worker nodes use \textbf{Apache Zookeeper} to store their states. The actual processing is done by the \textbf{worker nodes}, and each worker node runs multiple Worker Processes each of them mapped to their own topology \cite{stormtwitter}. It is also worth while noting that multiple worker processes in a worker node could be executing different parts of the topology.

\subsubsection{Worker Node}
\paragraph{Worker Node} is responsible for all the heavy lifting required for processing of the topology. One worker node may run a multiple number of worker processes. Each of these worker processes may belong to the same topology or different topologies. A worker node maintains a \textbf{Supervisor} and a multiple number of Worker Processes which in turn maintains a multiple number of executors. Thus, Executors enable intra-topology parallelism whereas Tasks provide intra-spout or intra-bolt parallelism. Figure \ref{fig:workerarch} depicts the design of a Worker node.\cite{stormtwitter}

@@

\paragraph{Supervisor} is running in every Worker Node and communicates with the Nimbus to receive assignments from it. Apart from taking assignments from the Nimbus it is also responsible for monitoring the health of the of the worker processes and respawns the worker processes in case it dies.

@@

\paragraph{Executors} are made up of one or multiple Tasks. Unless specified explicitly, storm assigns single task to a single executor. Executors takes tuples from the in queue, examines the task this tuple is destined to, performs the task and then places the result in the out queue \cite{stormtwitter}.

\subsection{Topology}
\paragraph{Topology} is a directed graph where the vertices represent computation and the edges represents the data flow. Another important thing about topologies in Storm is that they are allowed to have cycles. Figure \ref{fig:topo} shows a basic Storm Topology, which consists of Spouts and Bolts. Spouts are the Sources of tuples in a Topology whereas Bolts are the units which process these generated tuples. The user first submits the topology to the Apache Storm Nimbus, which consults with each of the available Supervisors in the worker nodes and assigns each of them with tasks related to the topology.

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=0.85\textwidth, height= 0.23\textheight]{topo.png}
    \caption{Generic Storm Topology}
    \label{fig:topo}
   \end{center}
\end{figure}

@@

As we can see in figure \ref{fig:throughput}, an experiment conducted in \cite{stormtwitter} to check the resilience and performance of storm under loss of worker nodes. The worker nodes were turned off one by one in fixed durations, and the throughput of the system was measured at every instance. The durations and number of machines are shown below in table \ref{tab:experiment}. As seen, 3 machines were killed every 3 minutes, but the number of workers were limited to 50 at every instance, thus increasing the number of workers per machine. The graph \ref{fig:throughput} suggests that every 15 minutes, when the worker nodes were killed, the throughput was reduced significantly while storm tries to balance the topology again. The balancing does not require extensible amount of time and results into a new maximum throughput. As we can observe in the graph, the throughput decreses every time a node is removed, which is an expected phenomenon, due to the decrease in the computational capacity.

@@

Apache Storm supports two types of processing schemes, \textbf{at least once} where each tuple is processed as the name suggests, at least once, and the other \textbf{at most once}, where the tuples are either processed once or dropped in cases of failure. To provide \textbf{at least once} semantics, Storm maintains an acker bolt, which keeps track of the tuple flow in the directed acyclic graph.

@@

	\section{Stream Processing: Design Considerations}
A stream is a continuous flow of data from one point to the another. A Stream Processing system works on these data streams, the work may just be simple transformations, but also could be as complex as predictive analysis. A stream processing system depending on the nature of application has different requirements.

@@

\subsection{System Architecture}
As discussed above, a data processing stream has different requirements and according to these various requirements, the System can be modified to suit the needs. The major two System architectures currently available are the Lambda Architecture and the Lambda-less (or Kappa) Architecture.

@@

\subsection{Topology Design Considerations}
A storm topology design with the spouts and bolts and affects the overall system thus should be properly considered. A spout should inject data into the topology from external sources where as the bolt should be a smallest implementable task. A storm topology should also be considered with various considerations in mind:

\subsubsection{Scalability:} A designed topology should be considered with scalability in mind. The major question is what configuration of bolts enables us to process greater amount of data at real time. Let us consider a simple example of word count topology which counts the repetition of words and ranks the most used words. It can be performed by simply including a spout and two bolts, the first of which to count the number of words and the second to aggregate and rank it (see fig. \ref{fig:ranking} Topology Setup 1).
Since we should have only one instance of Aggregator, to be able to get the aggregated result, in cases of increased data load, the number of rolling count bolts can be increased, but the Aggregation Bolt cannot be increases thus adding increased load at this node. To relax the demand on this node, we can introduce another bolt, the Intermediate Ranking Bolt, which is responsible for ranking the words at a certain predefined intervals and passes the intermediate ranks to the aggregator. This reduces load on the Aggregator since we can increase the intermediate ranking bolts to cope with the increased load.

@@

\subsubsection{Resilience:} Apache storm includes APIs which allow users to develop solutions which are resilient in nature. Using the acker bolt as show in figure \ref{fig:acker}, users can track the progress of the tuple injected into the system. Thus, writing a spout which can replay the tuple in case of failures allows us to guarantee the processing of data. For additional details please follow Section \ref{sec:resilience}.

@@

\subsubsection{Description}
\paragraph{Let us consider} a company which produces a range of products, for example \textbf{Apple Inc.} produces IPhones, IPods, MacBooks, etc. For such a big company an understanding of the sentiment of the people is important to maintain a positive outlook. The company can also analyze the feedback people place on a specific product, to change and adjust the pricing schemes in real time to the demand.  At present there are many social media networks available and people tend to share their views on these networks,  \textbf{Twitter} is also one of these social networks where a number of posts are available publicly and can be analyzed to obtain a generalized view about a specific product or the company in general.

@@

\paragraph{This} is not the only possible solution, there are multiple other tools such as \textbf{Spark Streaming}, \textbf{Samza}, etc. but this paper considers Apache Storm as it is in the spotlight of this paper, for further details regarding other available tools and their comparisons with Apache Storm, please consider the section \ref{sec:othertools} for further details.

\subsection{Practical Implementation}
The implementation of this solution concluded in three sub phases, namely Design, Development and Deployment. In the design phase, the storm topology was thoroughly planned, specifically the spouts, bolts and how they are interconnected. In the development phase, these spouts, bolts and other required aspects of the software was programmed and lastly was deployed on a locally hosted storm server.

@@

To Process the stream of tweets, we need to parse these tweets injected into the topology. The ParseTweetBolt processes these tweets and then emits whether the tweet was positive or negative under the simplified scenario. These separate positive or negative remark on a tweet does not provide useful information unless it is aggregated, thus a ResultAggregationBolt is responsible for aggregating all the results of the Parser bolt and providing the result to the target in real time.

@@

\subsection{Deployment}
\paragraph{Deploying} a storm topology requires a running zookeeper, nimbus and at least one worker node. In this practical implementation we use a single node to handle all the processing tasks. The zookeeper as well as the nimbus was instantiated in the local machine, along with a supervisor for the worker node. The Storm User interface presents all these information in a comprehensible manner within a web browser, which can be observed in figure \ref{fig:stormui} The java code was compiled and packaged as a jar and submitted to nimbus for deployment (refer GitHub readme for further details). It was then deployed in storm, which can be observed in the Screen shot for the user interface in figure \ref{fig:stormui}.

@@

\section{Other Tools}
 \label{sec:othertools}
Stream processing is a topic under constant focus at present and a lot of tools as well as models have been designed and created to support it. This section makes a comparison of the available widely accepted tool today. The available tools are mentioned below in brief such as:

 \begin{enumerate}
 \item \textbf{Spark Streaming} is an Open Source based project for stream processing.  It relies on cluster computing systems and operates on batches. The batch window in Spark Streaming is comparatively small thus allowing stream processing.It also provides flexibility to work in different programming languages such as Java, Scala and Python.
 \item \textbf{Apache Samza} is an Open Source project for near realtime distributed stream processing framework. It uses Apache Kafka for messaging and Apache Hadoop YARN for fault tolerance, processor isolation, security and resource management.\footnote{http://samza.apache.org/}
 \item \textbf{Flink Streaming} is also a Real time stream processing tool for distributed systems and builds batch processing on top of the Streaming engine.
 \end{enumerate}

@@

 \subsubsection{Development and Testing}
All three tools provide their own API for development and testing. A comparision among these tools can be made on a basis of the programming languages supported, Apache Samza currently only supports JVM languages, whereas Spark Streaming supports Java, Scala and Python. Apache Storm on the other hand supports JVM as well as Non JVM languages which increases development speed, as the developers can choose their language according to their comfort. Apart from the fact that choosing a language for development, users can utilize existing code, thus allowing high portability.

 \subsubsection{Tool setup}
Setting up of the tool in Apache Storm required minimum effort, achieved by downloading the zookeeper and storm binaries and unpacking them onto the computer. A minor configuration changes required attention, such as the location where zookeeper and storm would store data in. Apart from these details, it worked perfectly out of the box. Along the same line, Spark Streaming follows a similar approach for tool setup and can be installed by simply downloading the tarball and unpacking it. Apache Samza on the other hand maintains a repository \url(https://github.com/apache/samza-hello-samza) which contains the scripts which installs and starts the requirement dependencies in the host.

@@

 \subsubsection{Fault Tolerance}
As stream processing systems run for extended periods, it is quite necessary that these should be able to handle various kinds of faults and still keep processing data as they enter the system. The fault tolerance model for Storm has been extensively described in section \ref{sec:resilience}. Similarly, Spark Streaming and Apache Samza also use various techniques to resolve issues while deployed. Spark Streaming uses write ahead logs(journals) to keep track of all the data operation, and during failure of a worker, the worker is restarted and a state restore is done by reading the logs. This ensures that the data is processed as required. In comparison to Apache Storm, Storm continues to process incoming tuples even when the master node fails, but in Spark, the worker nodes stop working when the master node fails until a new leader starts. This is an added bonus to Storm, since a failed master manual restart would reinstate the state of the system without many changes.

@@

 \subsubsection{Benchmarks}
There have not been a comparative benchmark between Apache Storm, Spark Streaming and Apache Samza, but various sources have different comparisions between tools.  \cite{benchmark} considers Apache Storm, Spark Streaming and Apache Flink. The benchmark design consisted of counting the number of events taking place in the stream. The design was setup to read data from a Kafka Event Queue, the data being processed and result updated into Redis. The result included the window start time and the window end time, thus giving  an idea about the total latency of processing. According to this benchmark, they deduced that Apache Storm and Flink had low latency, whereas Spark Streaming had higher maximum throughput depending upon the batch duration setting. Figure \ref{fig:comparisiongraph} shows the comparative analysis among Apache Flink, Spark Streaming and Apache Storm. The batch processing scheme of Spark Streaming allows it to process events with higher throughput but with a trade off that the processing is delayed with higher latency. In the figure, we can observe that storm has comparatively less delay in processing, but according to \cite{benchmark}, storm with acking beyond 135,000 events per second, could not keep up with the throughput.

>>>>

Data can be considered the oil of this present age, because with huge amounts of data one can determine hidden patterns. A lot of companies utilize pattern recognition on big data to open opportunities for revenue generation whereas research institutions use it to discover patterns simply not observable by the human mind(simply due to the volume of data). The question on how to classify any data as big data can be answered by using the 3Vs: Volume, Variety and Velocity \cite{usingapachestorm}. Any data with big volume or variety (such as text, video, images, etc.) or that gets generated rapidly can be considered as big data. In this paper regarding \textbf{Apache Storm} we consider the third V of big data to be the influencing factor.

\textbf{Data Stream} is a sequence of data flowing from a source to a destination, for example a sensor reporting the temperature to a server. The sensor measures the temperature periodically and updates it to the server. Another example could also be the purchase requests of products in a shopping site. In all of these scenarios, the data is continuously being generated. Traditionally analyzing these types of data requires storage of these data onto persistence and later batch processing it to discern patterns as well as other useful information. But, with the help of \textbf{Apache Storm} we are able to process these data as they are generated with minimum latency.

@@

\paragraph{The Master node} also called the \textbf{Nimbus} is also responsible for the distribution and coordination of the execution of topologies. The Nimbus is the first point where the user submits the defined topology, which is then distributed among the worker nodes.  Figure \ref{fig:arch} illustrates the high level architecture where, Nimbus monitors all the available worker nodes and both the nimbus as well as the worker nodes use \textbf{Apache Zookeeper} to store their states. The actual processing is done by the \textbf{worker nodes}, and each worker node runs multiple Worker Processes each of them mapped to their own topology \cite{stormtwitter}. It is also worth noting that multiple worker processes in a worker node could be executing different parts of the topology.

\subsubsection{Worker Node}
\paragraph{Worker Node} is responsible for all the heavy lifting required for processing the topology. One worker node may run a multiple number of worker processes. Each of these worker processes may belong to the same topology or different topologies. A worker node maintains a \textbf{Supervisor} and a multiple number of Worker Processes which in turn maintains a multiple number of executors. Thus, Executors enable intra-topology parallelism whereas Tasks provide intra-spout or intra-bolt parallelism. Figure \ref{fig:workerarch} depicts the design of a Worker node.\cite{stormtwitter}

@@

\paragraph{Supervisor} is running in every Worker Node and communicates with the Nimbus to receive assignments from it. Apart from taking assignments from the Nimbus it is also responsible for monitoring the health of the worker processes and respawns the worker processes in case it dies.

@@

\paragraph{Executors} are made up of one or multiple Tasks. Unless specified explicitly, storm assigns single task to a single executor. Executors take tuples from the in queue, examines the task this tuple is destined to, performs the task and then places the result in the out queue \cite{stormtwitter}.

\subsection{Topology}
\paragraph{Topology} is a directed graph where the vertices represent computation and the edges represent the data flow. Another important thing about topologies in Storm is that they are allowed to have cycles. Figure \ref{fig:topo} shows a basic Storm Topology, which consists of Spouts and Bolts. Spouts are the Sources of tuples in a Topology whereas Bolts are the units which process these generated tuples. The user first submits the topology to the Apache Storm Nimbus, which consults with each of the available Supervisors in the worker nodes and assigns each of them with tasks related to the topology.

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=0.85\textwidth]{topo.png}
    \caption{Generic Storm Topology}
    \label{fig:topo}
   \end{center}
\end{figure}

@@

As we can see in figure \ref{fig:throughput}, an experiment conducted in \cite{stormtwitter} to check the resilience and performance of storm under loss of worker nodes. The worker nodes were turned off one by one in fixed durations, and the throughput of the system was measured at every instance. The durations and number of machines are shown below in table \ref{tab:experiment}. As seen, 3 machines were killed every 3 minutes, but the number of workers were limited to 50 at every instance, thus increasing the number of workers per machine. The graph \ref{fig:throughput} suggests that every 15 minutes, when the worker nodes were killed, the throughput was reduced significantly while storm tries to balance the topology again. The balancing does not require extensible amount of time and results into a new maximum throughput. As we can observe in the graph, the throughput decreases every time a node is removed, which is an expected phenomenon, due to the decrease in the computational capacity.

@@

Apache Storm supports two types of processing schemes, \textbf{at least once} where each tuple is processed as the name suggests, and the other \textbf{at most once}, where the tuples are either processed once or dropped in cases of failure. To provide \textbf{at least once} semantics, Storm maintains an acker bolt, which keeps track of the tuple flow in the directed acyclic graph.

@@

	\section{Stream Processing: Design Considerations}
A stream is a continuous flow of data from one point to the other. A Stream Processing system works on these data streams, the work may just be simple transformations, but could also be as complex as predictive analysis. A stream processing system depending on the nature of application has different requirements.

@@

\subsection{System Architecture}
As discussed above, a data processing stream has different requirements and according to these various requirements, the System can be modified to suit these needs. The major two System architectures currently available are the Lambda Architecture and the Lambda-less (or Kappa) Architecture.

@@

\subsection{Topology Design Considerations}
A storm topology design with the spouts and bolts and affects the overall system thus should be properly considered. A spout should inject data into the topology from external sources whereas the bolt should be a smallest implementable task. A storm topology should also be considered with various considerations in mind:

\subsubsection{Scalability:} A designed topology should be considered with scalability in mind. The major question is what configuration of bolts enables us to process greater amount of data at real time. Let us consider a simple example of word count topology which counts the repetition of words and ranks the most used words. It can be performed by simply including a spout and two bolts, the first of which to count the number of words and the second to aggregate and rank it (see fig. \ref{fig:ranking} Topology Setup 1).
Since we should have only one instance of Aggregator, to be able to get the aggregated result, in cases of increased data load, the number of rolling count bolts can be increased, but the Aggregation Bolt cannot be increases thus adding increased load at this node. To relax the demand on this node, we can introduce another bolt, the Intermediate Ranking Bolt, which is responsible for ranking the words at certain predefined intervals and passes the intermediate ranks to the aggregator. This reduces load on the Aggregator since we can increase the intermediate ranking bolts to cope with the increased load.

@@

\subsubsection{Resilience:} Apache storm includes APIs which allow users to develop solutions which are resilient in nature. Using the acker bolt as shown in figure \ref{fig:acker}, users can track the progress of the tuple injected into the system. Thus, writing a spout which can replay the tuple in case of failures allows us to guarantee the processing of data. For additional details please follow Section \ref{sec:resilience}.

@@

\subsubsection{Description}
\paragraph{Let us consider} a company which produces a range of products, for example \textbf{Apple Inc.} produces IPhones, IPods, MacBooks, etc. For such a big company an understanding of the sentiment of the people is important to maintain a positive outlook. The company can also analyze the feedback people place on a specific product, to change and adjust the pricing schemes in real time to the demand.  At present there are many social media networks available and people tend to share their views on these networks,  \textbf{Twitter}, for instance, is also one of these social networks where a number of posts are available publicly and can be analyzed to obtain a generalized view about a specific product or the company in general.

@@

\paragraph{This} is not the only possible solution, there are multiple other tools such as \textbf{Spark Streaming}, \textbf{Samza}, etc. but this paper concentrates Apache Storm as it is in the spotlight of this paper, for further details regarding other available tools and their comparisons with Apache Storm, please consider the section \ref{sec:othertools} for further details.

\subsection{Practical Implementation}
The implementation of this solution concluded in three sub phases, namely Design, Development and Deployment. In the design phase, the storm topology was thoroughly planned, specifically the spouts, bolts and how they are interconnected. In the development phase, these spouts, bolts and other required aspects of the software were programmed and lastly deployed on a locally hosted storm server.

@@

To Process the stream of tweets, we need to parse these tweets injected into the topology. The ParseTweetBolt processes these tweets and then emits whether the tweet was positive or negative under the simplified scenario. This separate positive or negative remark on a tweet does not provide useful information unless it is aggregated, thus a ResultAggregationBolt is responsible for aggregating all the results of the Parser bolt and providing the result to the target in real time.

@@

\subsection{Deployment}
\paragraph{Deploying} a storm topology requires a running zookeeper, nimbus and at least one worker node. In this practical implementation we use a single node to handle all the processing tasks. The zookeeper as well as the nimbus was instantiated in the local machine, along with a supervisor for the worker node. The Storm User interface presents all these information in a comprehensible manner within a web browser, which can be observed in figure \ref{fig:stormui}, the java code was compiled and packaged as a jar and submitted to nimbus for deployment (refer GitHub readme for further details). It was then deployed in storm, which can be observed in the Screen shot for the user interface in figure \ref{fig:stormui}.

@@

\section{Other Tools}
 \label{sec:othertools}
Stream processing is a topic under constant focus at present and a lot of tools as well as models have been designed and created to support it. This section makes a comparison of the available widely accepted tools today. The available tools are mentioned below in brief such as:

 \begin{enumerate}
 \item \textbf{Spark Streaming} is an Open Source based project for stream processing.  It relies on cluster computing systems and operates on batches. The batch window in Spark Streaming is comparatively small thus allowing stream processing. It also provides flexibility to work in different programming languages such as Java, Scala and Python.
 \item \textbf{Apache Samza} is an Open Source project for near realtime distributed stream processing framework. It uses Apache Kafka for messaging and Apache Hadoop YARN for fault tolerance, processor isolation, security and resource management.\footnote{http://samza.apache.org/}
 \item \textbf{Flink Streaming} is also a Real time stream processing tool for distributed systems and builds batch processing on top of the Streaming engine.
 \end{enumerate}

@@

 \subsubsection{Development and Testing}
All three tools provide their own API for development and testing. A comparision among these tools can be made on a basis of the programming languages supported, Apache Samza currently only supports JVM languages, whereas Spark Streaming supports Java, Scala and Python. Apache Storm on the other hand supports JVM as well as Non JVM languages which increases development speed, as the developers can choose their language according to their comfort. Apart from being able to choose a language for development, users can utilize existing code, thus allowing high portability.

 \subsubsection{Tool setup}
Setting up of the tool in Apache Storm requires minimum effort, achieved by downloading the zookeeper and storm binaries and unpacking them onto the computer. A minor configuration changes required attention, such as the location where zookeeper and storm would store data in. Apart from these details, it works perfectly out of the box. Along the same line, Spark Streaming follows a similar approach for tool setup and can be installed by simply downloading the tarball and unpacking it. Apache Samza on the other hand maintains a repository \url(https://github.com/apache/samza-hello-samza) which contains the scripts which installs and starts the requirement dependencies in the host.

@@

 \subsubsection{Fault Tolerance}
As stream processing systems run for extended periods, it is quite necessary that these are able to handle various kinds of faults and still keep processing data as they enter the system. The fault tolerance model for Storm has been extensively described in section \ref{sec:resilience}. Similarly, Spark Streaming and Apache Samza also use various techniques to resolve issues while deployed. Spark Streaming uses write ahead logs(journals) to keep track of all the data operation, and during failure of a worker, the worker is restarted and a state restore is done by reading the logs. This ensures that the data is processed as required. In comparison to Apache Storm, Storm continues to process incoming tuples even when the master node fails, but in Spark, the worker nodes stop working when the master node fails until a new leader starts. This is an added bonus to Storm, since a failed master manual restart would reinstate the state of the system without many changes.

@@

 \subsubsection{Benchmarks}
There has not been a comparative benchmark between Apache Storm, Spark Streaming and Apache Samza, but various sources have different comparisions between tools.  \cite{benchmark} considers Apache Storm, Spark Streaming and Apache Flink. The benchmark design consisted of counting the number of events taking place in the stream. The design was setup to read data from a Kafka Event Queue, the data being processed and result updated into Redis. The result included the window start time and the window end time, thus giving  an idea about the total latency of processing. According to this benchmark, they deduced that Apache Storm and Flink had low latency, whereas Spark Streaming had higher maximum throughput depending upon the batch duration setting. Figure \ref{fig:comparisiongraph} shows the comparative analysis among Apache Flink, Spark Streaming and Apache Storm. The batch processing scheme of Spark Streaming allows it to process events with higher throughput but with a trade off that the processing is delayed with higher latency. In the figure, we can observe that storm has comparatively less delay in processing, but according to \cite{benchmark}, storm with acking beyond 135,000 events per second, could not keep up with the throughput.
